Title here

### 2016-12-06 portfolio/text/ms.Rmd
Add Eric's edits

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, otherwise well-intentioned regulations can limit
the actions available to individuals to reduce income variability, or
incentivize against diversification. For example, in terrestrial agriculture,
crop subsidies in the US may incentivize some farms to specialize (e.g. dairy,
grains, soybeans; @mishra2002). In other instances, external market forces
(prices, demand) may discourage the adoption of diverse strategies and help
promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
available to commercial fishers [@kasperski2013] and both fish and fishers are
spatially mobile. However, fishers still have the ability to lower risk. They
can, for example, target multiple species as permitted by a license, own
multiple licenses, or diversify spatially and fish in multiple regions. The
effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization (REFS). The majority of papers that examine effects of
diversification on revenue do so at aggregate levels (e.g. households, counties,
communities; REFS), rather than focus on individuals. Because of this, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
many permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. We develop and apply a hierarchical
Bayesian model to understand the effect of diversity on both individual revenue
and revenue variability that builds on results from vessel [@kasperski2013],
fleet [@sethi2012b], and community-level [@sethi2014] variability. Our model
controls for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

@@

We developed a hierarchical variance-function regression model [@davidian1987;
@hall1989], which separately models changes in the mean and variance of
individual revenue as a function of covariates. Whereas traditional
hierarchical linear models assume homoscedasticity --- that the residual
variance does not vary systematically --- the variance-function regression model
allows the variance to be heteroscedastic and vary as a function of predictors.
Our specific model allows certain fishing strategies and levels of species
diversification to generate more or less revenue and be more or less variable
than others.

@@

We combined similar permits into permit groupings and identified sets of permit
groupings held by an individual in a specific year as "strategies" (Supporting
Information). For example, we combined permits that were the same except for
vessel size (e.g. a longliner fishing sablefish on a vessel under 60 ft was
considered to have the same strategy as a longliner fishing sablefish on
a vessel over 60 ft), and we combined a number of permits that were targeting
a single species and differed only in region fished. Therefore, for example,
someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska
Peninsula were both considered to have a 'herring roe' permit group. For salmon
permits, we grouped permits using the same gear type across regions for groups
that shared similar species composition. After this process of combining
permits, we were left with `r n_grouped_permits` unique fishing permit groups
(Table S\ref{tab:strategies}), which we aggregated within each person-year
combination to form permit strategies. For example, if someone fished halibut
and sablefish in the same year, their strategy for that year would be
"halibut-sablefish". Specific details on our strategy definitions are available
in the Supporting Information.

@@

\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
or effective season length as the sum of the season lengths for each permit each
year. Season length for a given permit was calculated as the number of days
between the first instance of fishing a permit and the last instance of fishing
a permit each year. For example, if someone fished for 30 days on a salmon drift
gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut
permit, that individual's season length would be 70 days.

@@

We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ parameters to allow for heavier
tails. We ran four chains and 2,000 iterations and discarded the first 1,000
iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
chain convergence visually with trace plots. We also ensured that $\hat{R}<1.05$
(the potential scale reduction factor) and that the effective sample size, as
calculated by the rstan R package [@standevelopmentteam2015], was greater than
200 for all parameters [@gelman2014]. We verified that our model returned
sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
with the lme4 R package [@bates2015], and by fitting our model to simulated data
and recovering the true values (Supporting Information). Code to reproduce our
results (with a simulated data set) is available at
<https://github.com/NCEAS/pfx-commercial>.

>>>>

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, external forces may restrict the options available
to individuals to reduce variability. For example, otherwise well-intentioned
regulations may limit how individuals diversify, or incentivize against
diversification. For example, in terrestrial agriculture, crop subsidies in the
US may incentivize some farms to specialize (e.g. dairy, grains, soybeans;
@mishra2002). In other instances, external market forces (prices, demand) may
discourage the adoption of diverse strategies and help promote specialization
[@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) have not been
obtainable by commercial fishers [@kasperski2013], possibly because both fish
and fishers are spatially mobile. However, fishers still have the ability to
lower risk. They can, for example, target as many species as permitted by
a license, they may own multiple licenses, or diversify spatially and fish in
multiple regions (increasing their total catch, or numbers of species caught).
The effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization for both farmers and fishers [@abdulai2000; @arnason2005]. The
majority of papers that examine effects of diversification on revenue do so at
aggregate levels that combine individuals (e.g. households, counties,
communities; @sethi2014; @kasperski2013). Because of this, key uncertainties
remain about the fundamental relationship between diversity, fishing revenue,
and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
many permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit the
entry of new participants through individual fishing quotas (IFQs) [@grimm2012].

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. Similarly, we are able to disentagnle
fishers that share the same vessel. We develop and apply a hierarchical Bayesian
model to understand the effect of diversity on both individual revenue and
revenue variability that builds on results from vessel [@kasperski2013], fleet
[@sethi2012b], and community-level [@sethi2014] variability. Our model controls
for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

@@

We developed a hierarchical variance-function regression model [@davidian1987;
@hall1989], which separately but simultaneously models changes in the mean and variance of
individual revenue as a function of covariates. Whereas traditional
hierarchical linear models assume homoscedasticity --- that the residual
variance does not vary systematically --- the variance-function regression model
allows the variance to be heteroscedastic and vary as a function of predictors.
Our specific model allows certain fishing strategies and levels of species
diversification to generate more or less revenue and be more or less variable
than others.

@@

We used combinations of state and federal fishing permits held by an individual
in in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a 'herring
roe' permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species compositon (Supporting Information).

After this process of combining
permits, we were left with `r n_grouped_permits` unique fishing permit groups
(Table S\ref{tab:strategies}), which we aggregated within each person-year
combination to form permit strategies. For example, if someone fished halibut
and sablefish in the same year, their strategy for that year would be
"halibut-sablefish". Specific details on our strategy definitions are available
in the Supporting Information.

@@

\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
or effective season length as the sum of the season lengths for each permit
each year. Season length for a given permit was calculated as the number of
days between the first instance of fishing a permit and the last instance of
fishing a permit each year. For example, if someone fished for 30 days on a
salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
longline halibut permit, that individual's season length was calculated as 70
days.

@@

We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ parameters to allow for heavier
tails. We ran four chains and 2,000 iterations and discarded the first 1,000
iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
chain convergence visually with trace plots. We also ensured that
$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
sample size, as calculated by the rstan R package [@standevelopmentteam2015],
was greater than 200 for all parameters [@gelman2014]. We verified that our
model returned sensible estimates by comparing our estimates to an ad-hoc
two-stage model fit with the lme4 R package [@bates2015], where we first fit a
mixed effect regression model to the mean, and secondly modeled the absolute
residuals from that first model in a second mixed effects model representing
the variance component. We also validated the model with simulated data, and
were able to recover the true values (Supporting Information). Code to
reproduce our results (with a simulated data set) is available at
<https://github.com/NCEAS/pfx-commercial>.
